---
title: "ImpExp-Colombia"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Pre-work
First we need to access to our dataset and extract the available files, import the necessary libraries and check if they are updated. 
####Install packages
```{r}
install.packages("rmarkdown")
install.packages("readr")
install.packages("dplyr")
install.packages("plyr")
```

####Activating packages
```{r}
library(rmarkdown)
library(readr)
library(dplyr)
library(plyr)
```
---


##Importing Data
This is what we have:

*   **colombia_exports.csv** [16.7MB] and **colombia_imports.csv** [24.1MB]: with historical data from colombian international market
*   **country_names.csv** [6.5kB]: with the list of official abbreviation of the countries in the study
*   **groups_sitc_rev2.csv** [2.8kB]: with the category groups of products from the SITC
*   **products_sitc_rev2.csv** [39.4kB]: with detailed products and the official ID from the SITC

Now we can import the five available files separated by "|" but in csv format:
```{r}
country_names <- read_delim("data/country_names.csv", 
                            "|", escape_double = FALSE, na = "NA", 
                            trim_ws = TRUE)

colombia_imports <- read_delim("data/colombia_imports.csv", 
                               "|", escape_double = FALSE, na = "NA", 
                               trim_ws = TRUE)

colombia_exports <- read_delim("data/colombia_exports.csv", 
                               "|", escape_double = FALSE, na = "NA", 
                               trim_ws = TRUE)

groups_sitc_rev2 <- read_delim("data/groups_sitc_rev2.csv", 
                               "|", escape_double = FALSE, na = "NA", 
                               trim_ws = TRUE)

products_sitc_rev2 <- read_delim("data/products_sitc_rev2.csv", 
                                 "|", escape_double = FALSE, na = "NA", 
                                 trim_ws = TRUE)
```
---


##Cleaning and Transforming Data
We can take a quick look at the components of each DataFrame using the functions `head()` to see the first five rows, `str()` to see the number of rows and columns that make it up and `sum(is.nan())` to check if there are empty cells that can generate an error. 

But, in resume:
*   **colombia_exports** have 7 variables and 389.596 observations. No Na's
*   **colombia_imports** have 7 variables and 560.492 observations. No Na's
*   **country_names** have 4 variables and 263 observations. No Na's
*   **groups_sitc_rev2** have 3 variables and 65 observations. No Na's
*   **products_sitc_rev2** have 4 variables and 988 observations. No Na's


####Adding tags and merge in one table
```{r}
colombia_imports <- mutate(colombia_imports, category="import")
colombia_exports <- mutate(colombia_exports, category="export")
impExp_Colombia <- rbind(colombia_imports, colombia_exports) 
```


####Renaming columns
```{r}
names(impExp_Colombia) <- c("Index", "Unnamed: 0", "Year", "Origin", "Destination", "Product", "Value US$", "Category")
```